---
title: "PM10 and PM2.5 Analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Required Libraries

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(ggprism)  # Remove if not available
```

## PM10 Analysis

### Load and Prepare Data

```{r}
file_path <- "pm10_data.csv"  # Update with the actual file path
df <- read_csv(file_path, skip = 1, col_types = cols())

colnames(df) <- c("Date", "Source", "Site_ID", "POC", "Daily_Mean_PM10", "Units", "Daily_AQI_Value", "Local_Site_Name", 
                  "Daily_Obs_Count", "Percent_Complete", "AQS_Parameter_Code", "AQS_Parameter_Description", 
                  "Method_Code", "CBSA_Code", "CBSA_Name", "State_FIPS_Code", "State", "County_FIPS_Code", "County", 
                  "Site_Latitude", "Site_Longitude")

df$Date <- as.Date(df$Date, format="%m/%d/%Y")
df$Year <- format(df$Date, "%Y")
df$Month <- format(df$Date, "%m")
df$Daily_Mean_PM10 <- as.numeric(df$Daily_Mean_PM10)
```

### Monthly Average PM10 Concentration

```{r}
df_grouped <- df %>%
  group_by(Year, Month) %>%
  summarise(Average_PM10 = mean(Daily_Mean_PM10, na.rm = TRUE))

df_grouped$Date <- as.Date(paste(df_grouped$Year, df_grouped$Month, "01", sep = "-"))
```

### PM10 Scatter Plot

```{r}
ggplot(df_grouped, aes(x = Date, y = Average_PM10)) +
  geom_point(color = "blue", size = 3) +
  geom_line(color = "red", linetype = "dashed") +
  labs(title = "Monthly Average PM10 Concentration", x = "Date", y = "PM10 Concentration (µg/m³)") +
  theme_prism()
```

### Save Processed Data

```{r}
write_csv(df_grouped, "monthly_avg_pm10.csv")
```

## PM2.5 Analysis

### Load and Prepare Data

```{r}
df <- read_csv(file_path, skip = 1, col_types = cols())

colnames(df) <- c("Date", "Source", "Site_ID", "POC", "Daily_Mean_PM2.5_Concentration", "Units", "Daily_AQI_Value", 
                  "Local_Site_Name", "Daily_Obs_Count", "Percent_Complete", "AQS_Parameter_Code", "AQS_Parameter_Description", 
                  "Method_Code", "CBSA_Code", "CBSA_Name", "State_FIPS_Code", "State", "County_FIPS_Code", "County", 
                  "Site_Latitude", "Site_Longitude")

df$Date <- as.Date(df$Date, format="%m/%d/%Y")
df$Year <- format(df$Date, "%Y")
df$Month <- format(df$Date, "%m")
df$Daily_Mean_PM2.5_Concentration <- as.numeric(df$Daily_Mean_PM2.5_Concentration)
```

### Monthly Average PM2.5 Concentration

```{r}
df_grouped <- df %>%
  group_by(Year, Month) %>%
  summarise(Average_PM2.5 = mean(Daily_Mean_PM2.5_Concentration, na.rm = TRUE))

df_grouped$Date <- as.Date(paste(df_grouped$Year, df_grouped$Month, "01", sep = "-"))
```

### PM2.5 Scatter Plot

```{r}
ggplot(df_grouped, aes(x = Date, y = Average_PM2.5)) +
  geom_point(color = "blue", size = 3) +
  geom_line(color = "red", linetype = "dashed") +
  labs(title = "Monthly Average PM2.5 Concentration", x = "Date", y = "PM2.5 Concentration (µg/m³)") +
  theme_prism()
```

### Save Processed Data

```{r}
write_csv(df_grouped, "monthly_avg_pm2.5.csv")
```

## Conclusion

This report presents an analysis of PM10 and PM2.5 concentrations using air quality monitoring data. The visualizations provide insights into monthly variations in particulate matter levels.
